---
title: "p8105_hw1_hh2767"
author: "Haoran Hu"
date: "2018-09-18"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r include = FALSE}
library(tidyverse)
```

In this part, I will construct a data frame with four columns. The four columns are: a random sample of size 10 from a uniform[0, 5] distribution, a logical vector indicating whether elements of the sample are greater than 2, a length-10 character vector, and a length-10 factor vector, from left to right.

```{r}
problem1_df = tibble(
  vec_unif = runif(10, min = 0, max = 5),
  vec_logical_compare_2 = ifelse(vec_unif > 2, TRUE, FALSE),
  vec_char = c("I", "am", "doing", "my", "second", "hw", "of", "p8015", "data", "science"),
  vec_factor = factor(c("male", "female", "male", "female", "male", "male",     "female", "male", "female", "male"))
  )
```
The data frame is shown below:
```{r echo = FALSE}
problem1_df
```


In this part, I will try to take the mean of each variable in the data frame. It only wroked for the first colomn(vec_nuif).