---
title: "p8105_hw1_hh2767"
author: "Haoran Hu"
date: "2018-09-18"
output: github_document
   
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r include = FALSE}
library(tidyverse)
```
#Problem 1
##Creating a data frame
In this part, I will construct a data frame with four columns. From left to right, the four columns are: a random sample of size 10 from a uniform [0, 5] distribution, a logical vector indicating whether elements of the sample are greater than 2, a length-10 character vector, and a length-10 factor vector.
```{r}
problem1_df = tibble(
  vec_unif = runif(10, min = 0, max = 5),
  vec_lgl_greater_2 = ifelse(vec_unif > 2, TRUE, FALSE),
  vec_char = c("I", "am", "doing", "my", "second", "hw", "of", "p8015", "data", "science"),
  vec_factor = factor(c("male", "female", "male", "female", "male", "male",     "female", "male", "female", "male"))
  )
```
The data frame is as below:
```{r echo = FALSE}
problem1_df
```
##Take the mean of each variable in the data frame
In this part, I will try to take the mean of each variable in the data frame. It only wroked for the first column(vec_nuif) and the second column(vec_lgl_greater_2).
```{r}
mean_vec_unif = mean(problem1_df$vec_unif)
mean_vec_lgl_greater_2 = mean(problem1_df$vec_lgl_greater_2)
```
The mean of "vec_unif" variable is `r mean_vec_unif`, and the mean of "vec_lgl_greater_2" variable is `r mean_vec_lgl_greater_2`. The mean of "vec_unif" is directly calculated, while the mean of "vec_lgl_greater_2" was calculated by treating "TRUE" as numerical "1", and "FALSE" as numerical "2". The mean of "vec_char" and "vec_factor" cannot be calculated, because they are neither numerical value nor lgl value, and if we calculate the mean of them, the outputs will be displayed as NAs.

##Convert variables from one type to another
In this part, I will apply the "as.numeric" function to the logical, character, and factor variables, but the output will not be shown.
```{r eval = FALSE}
as.numeric(problem1_df$vec_lgl_greater_2)
as.numeric(problem1_df$vec_char)
as.numeric(problem1_df$vec_factor)
```
In the following chunk, I will try converting "vec_char" from character to factor to numeric, and converting "vec_factor" from factor to character to numeric.

```{r}
problem1_df$vec_char = as.factor(problem1_df$vec_char) #converting vec_char from character to factor
problem1_df$vec_char = as.numeric(problem1_df$vec_char) #converting vec_char from factor to numeric
problem1_df$vec_factor = as.character(problem1_df$vec_factor) #converting vec_factor from factor to character
problem1_df$vec_factor = as.numeric(problem1_df$vec_factor) #converting vec_factor from character to numeric
```
The result is as below:
```{r echo = FALSE}
problem1_df
```
Therefore, we can successfully convert character to factor to numeric, but we cannot convert factor to character to numeric, because NAs are introduced by coercion when converting character to numeric.

#Problem 2
##Creating a data frame
The following chunk creates a data frame with 5 columns. From left to right, the columns are a random sample of size 1000 from a standard Normal distribution, a random sample of size 1000 from a standard Normal distribution, a logical vector indicating whether the x + y > 0, a numeric vector created by coercing the above logical vector, and a factor vector created by coercing the above logical vector, respectively. 
```{r}
problem2_df = tibble(
  x = rnorm(1000, mean = 0, sd = 1),
  y = rnorm(1000, mean = 0, sd = 1),
  vec_lgl_xy_greater_0 = ifelse((x + y) > 0, TRUE, FALSE),
  vec_lgl_to_num = as.numeric(vec_lgl_xy_greater_0),
  vec_lgl_to_fact = as.factor(vec_lgl_xy_greater_0)
  )
problem2_df
```
The data frame is a `r nrow(problem2_df)` * `r ncol(problem2_df)` data set. The mean of x is `r mean(problem2_df$x)`, and the median of x is `r median(problem2_df$x)`.The proportion of cases for which the logical vector is TRUE is `r mean(problem2_df$vec_lgl_xy_greater_0)`.

##Scatterplot colored in different ways
A scatterplot that color points using the logical variable is as below. The plot is exported to the project directory using ggsave.
```{r}
ggplot(problem2_df, aes(x = x, y = y, color = vec_lgl_xy_greater_0)) + geom_point()
ggsave("scatplot_bylgl.pdf", height = 4, width = 6, path = ".\\plot")
```
A scatterplot that color points using the numeric variable is as below.
```{r}
ggplot(problem2_df, aes(x = x, y = y, color = vec_lgl_to_num)) + geom_point()
```
A scatterplot that color points using the factor variable is as below.
```{r}
ggplot(problem2_df, aes(x = x, y = y, color = vec_lgl_to_fact)) + geom_point()
```
As the plots show, when color points using logical or factor vectors, the color scale is two discrete values: TRUE or FALSE; when color points using numerical vector, the color scale is [0,1].

